---
title: "JHU COVID-19 Data by State and National"
author: "Jonah Pool"
date: "6/8/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
update: 6/8/2020
---

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(flextable)
library(sjPlot)
library(lubridate)
library(DescTools)
library(effsize)
library(pwr)
library(lmerTest)
library(scales)
library(kableExtra)
library(knitr)
library(fpp2)

library(treemap)

#library(ggspatial)
#library(rnaturalearth)
#library(rnaturalearthdata)
```

Constants
```{r}
yd <- today()  - ddays(1)
td <- today()

td
day <- yd
MichiganRiots<- seq(ymd("2020-05-26"), ymd("2020-06-03"), by = 1)
exclude = c("")
# date for daily data release
a <- as.character(ymd(day))
b <- str_split(a, "-")
b1 <- as.character(b[[1]][2])
b2 <- as.character(b[[1]][3]) 
b3 <- as.character(b[[1]][1])
d_lookup_str <- str_c(b1, b2, b3, sep = "-")
sinceMarch16 <- seq(date("2020-03-16"), day, by=1)
#sinceMarch16
last30 <- seq(td-30, td, by=1)
last21 <- seq(td-21, td, by=1)
last14 <- seq(td-14, td, by=1)
last7 <- seq(td-7, td, by=1)

red_states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "Florida", "Georgia", "Idaho", "Indiana", "Iowa", "Maryland", "Massachusetts", "Mississippi", "Missouri", "Nebraska", "New Hampshire", "North Dakota", "Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "West Virginia", "Wyoming")

blue_states <- c("California", "Colorado", "Connecticut", "Delaware", "Hawaii", "Illinois", "Kansas", "Kentucky", "Louisiana", "Maine", "Michigan", "Minnesota", "Montana", "Nevada", "New Jersey", "New Mexico", "New York", "North Carolina", "Oregon", "Pennsylvania", "Rhode Island", "Virginia", "Washington", "Wisconsin")


bold.14.text <- element_text(face = "bold", size = 12)

red.fill.5 <- c("#bf080e","#bf080e","#bf080e","#bf080e","#bf080e")
blue.fill.5 <- c("#0016bd","#0016bd","#0016bd","#0016bd","#0016bd")

red.fill.10 <- c("#bf080e","#bf080e","#bf080e","#bf080e","#bf080e","#bf080e","#bf080e","#bf080e","#bf080e","#bf080e", "#bf080e")
blue.fill.10 <- c("#0016bd","#0016bd","#0016bd","#0016bd","#0016bd","#0016bd","#0016bd","#0016bd","#0016bd","#0016bd", "#0016bd")

x.limit.march15 <- seq(ymd("2020-03-15"), day, by=1)

#annotation.moving.average <- annotate(geom = "text", x = xrange[1], y = yrange[2], label = paste(strwrap("Note: 5 Day Moving Average", 40), collapse = "\n"), hjust = 0, vjust = 1, size = 4)

pq <- theme(text = bold.14.text) +
  #theme(axis.ticks = element_blank()) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=2),
        panel.background=element_rect(colour = NA, fill = "white"))
jhu_caption <- labs(caption="Data Source: Johns Hopkins University CSSE")
```

Functions
```{r}
movingAverage <- function(x, n=14, centered=TRUE) {
    
    if (centered) {
        before <- floor  ((n-1)/2)
        after  <- ceiling((n-1)/2)
    } else {
        before <- n-1
        after  <- 0
    }

    # Track the sum and count of number of non-NA items
    s     <- rep(0, length(x))
    count <- rep(0, length(x))
    
    # Add the centered data 
    new <- x
    # Add to count list wherever there isn't a 
    count <- count + !is.na(new)
    # Now replace NA_s with 0_s and add to total
    new[is.na(new)] <- 0
    s <- s + new
    
    # Add the data from before
    i <- 1
    while (i <= before) {
        # This is the vector with offset values to add
        new   <- c(rep(NA, i), x[1:(length(x)-i)])

        count <- count + !is.na(new)
        new[is.na(new)] <- 0
        s <- s + new
        
        i <- i+1
    }

    # Add the data from after
    i <- 1
    while (i <= after) {
        # This is the vector with offset values to add
        new   <- c(x[(i+1):length(x)], rep(NA, i))
       
        count <- count + !is.na(new)
        new[is.na(new)] <- 0
        s <- s + new
        
        i <- i+1
    }
    
    # return sum divided by count
    s/count
}

bar_chart_ts <- function(df, y, custom.fill = blue.fill.10, lab.title, y.lim = max(y)) {
  plot <- ggplot(df, aes(x = date, y = y)) + 
    geom_col() +
    scale_fill_manual(values = custom.fill) +
    labs(y = "", title = lab.title, x = "Date") +
    theme(legend.position = "none") +
    ylim(0,y.lim) +
    pq
  plot
}

line_chart_ts <- function(df, y, custom.fill = blue.fill.10, lab.title, y.lim = max(y)) {
  plot <- ggplot(df, aes(x = date, y = y)) + 
    geom_line(size = 1.2) +
    scale_fill_manual(values = custom.fill) +
    scale_color_manual(values = custom.fill)+
    labs(y = "", title = lab.title, x = "Date") +
    theme(legend.position = "none") +
    ylim(0,y.lim) +
    pq
  plot
}

line_chart_ma_ts <- function(df, y, custom.fill = blue.fill.10, lab.title, y.lim = max(y)) {
  plot <- ggplot(df, aes(x = date, y = y)) + 
    geom_smooth(size = 1.2) +
    scale_fill_manual(values = custom.fill) +
    scale_color_manual(values = custom.fill)+
    labs(y = "", title = lab.title, x = "Date") +
    theme(legend.position = "none") +
    ylim(0,y.lim) +
    pq
  plot
}

```

Load Johns Hopkins Univ. CSSE Data
```{r}
ts_date_count <- today() - ymd(20200122)

daily_fname <- str_c("../COVID-19/csse_covid_19_data/csse_covid_19_daily_reports_us/", d_lookup_str, ".csv")
csse_today <- read_csv(daily_fname, col_types = "ccTddddddddddddcdd")
csse_confirmed_us_ts <- read_csv("../COVID-19/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv",
                                 col_types=str_c('cccddcccddc', 
                                 str_c(rep("i", times=ts_date_count), collapse=""))
                                 )

csse_deaths_us_ts <- read_csv("../COVID-19/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv", 
                              col_types=str_c('cccddcccddc', 
                              str_c(rep("i", times=ts_date_count+1), collapse=""))
                              )
uid <- read_csv("../COVID-19/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv", col_types = "iccddcccddci")
daily_updates_ts <- read_csv("daily_updates_ts_us.csv", col_types="ccDdTdcdddddddddddddd")
```

16, 316, 630, 580, 850
Filter US ID Values
```{r}

us_uid <- uid %>% 
  filter(code3 %in% c(16, 316, 580, 850, 630, 840)) %>%
  select(UID, Admin2, Province_State, Combined_Key, Population) %>% 
  arrange(UID)
us_uid
states <- c(us_uid$Province_State[1:4], us_uid$Province_State[6:57])
states_uid <-seq(84000001, 84000056, by = 1)
uid_states <- us_uid %>% 
  filter(UID %in% c(seq(84000001, 84000056, by = 1))) %>% 
  filter(UID != 84000011)
nyc_uid <- uid$UID[uid$Combined_Key == "New York City, New York, US"]
nyc_pop <- uid$Population[uid$Combined_Key == "New York City, New York, US"]
```

By State Snapshot Today
```{r}

csse_today_states <- csse_today %>% 
  select(-Lat, -Long_, -FIPS, -Last_Update) %>%
  inner_join(uid_states, key = "UID") %>% 
  filter(Province_State %in% us_uid$Province_State & !(is.na(Province_State)), !(Province_State %in% c("Diamond Princess", "Recovered", "Grand Princess"))) %>%
  rename(state = Province_State) %>% 
  rename(confirmed = Confirmed, deaths = Deaths, recovered = Recovered, active = Active, tested = Total_Test_Results, hospitalized = People_Hospitalized) %>% 
  arrange(desc(Testing_Rate))
daily_updates_ts %<>% 
  select(-Lat, -Long_, -FIPS, -Last_Update) %>%
  inner_join(uid_states, key = "UID") %>% 
  #filter(Province_State %in% us_uid$Province_State & !(is.na(Province_State)), !(Province_State %in% c("Diamond Princess", "Recovered", "Grand Princess"))) %>%
  rename(state = Province_State) %>% 
  rename(date = Date, confirmed = Confirmed, deaths = Deaths, recovered = Recovered, active = Active, tested = People_Tested, hospitalized = People_Hospitalized) %>% 
  filter(!(is.na(Testing_Rate))) %>% 
  arrange(date)

```


Testing and Incident Rate
```{r}
csse_today_states %<>% 
  mutate(c_d_rat = deaths / confirmed) %>% 
  filter(!(is.na(c_d_rat))) %>% 
  filter(c_d_rat != Inf)

# shift data to days since 50th case and 10th recorded death

#tvi_plot <- daily_updates_ts %>% 
#  ggplot(aes(x = Testing_Rate, y = Hospitalization_Rate)) + 
#  geom_point(aes(group = state))+
#  geom_smooth(aes(group = state), method = 'lm')+
#  labs(y = "Hosp", title = strwrap("Effect of Testing Rate on Incident Rate of COVID-19"), subtitle = #"Within the USA and its territories", x = "Tested Individuals per 100,000 People", caption="Data Source: #Johns Hopkins University CSSE 2020-04-29.\n Each point represents one state/territory.") +
#  theme(legend.position = "right") +
#  scale_y_continuous()+
#  pq
#ggsave("testing-data-plot.png", plot=tvi_plot)
```

```{r}
#"1M", "2M", "3M", "4M", "5M", "6M", "7M", "8M", "9M", "10M", "11M", "12M"
md_testing_rate <- csse_today_states$Testing_Rate[csse_today_states$state == "Maryland"]
md_testing_rate
mean_testing_rate <- mean(csse_today_states$Testing_Rate)
mean_testing_rate

csse_today_states_sorted <- csse_today_states %>% mutate(vote = case_when(
  state %in% blue_states ~ "Dem",
  state %in% red_states ~ "Rep"
)) %>% 
  mutate(vote = factor(vote)) %>% mutate(vote = fct_relevel(vote, c("Rep", "Dem")))
head(daily_updates_ts)

in_millions <- function(x) (x / 1000000)
states_sorted_in_millions <- csse_today_states_sorted %>% mutate_at(vars(matches("Population")), in_millions) %>% pull(Population) %>% summary()
N <- 20
pop_breaks <- c(-Inf, 500000, seq(10^6,N*10^6, by = 2*10^6), Inf)
pop_break_levels <- c("<500K", "500K", paste(seq(2,N-2,by=2), rep("M",times = round(((N/2)-1), 1), sep="")), paste(c(">", as.character(N), "M"),collapse=""))
pop_break_levels

csse_today_states_sorted %<>% mutate(Pop_fac = cut(csse_today_states_sorted$Population, breaks = pop_breaks, labels = pop_break_levels))
csse_today_states_sorted %<>% filter(Testing_Rate < 75000)# %>% filter(Mortality_Rate < 6)

daily_updates_ts %<>% mutate(Pop_fac = cut(daily_updates_ts$Population, breaks = pop_breaks, labels = pop_break_levels))
daily_updates_ts %<>% filter(Testing_Rate < 75000)

mean_testing_rates <- csse_today_states_sorted %>% 
  group_by(vote) %>% 
  summarize(sum = sum(Population), avg = mean(Population), median = median(Population))
mean_testing_rates

daily_updates_ts_sorted <- daily_updates_ts %>% mutate(vote = case_when(
  state %in% blue_states ~ "Dem",
  state %in% red_states ~ "Rep"
)) %>% 
  mutate(vote = factor(vote)) %>% mutate(vote = fct_relevel(vote, c("Rep", "Dem")))
daily_updates_ts_sorted

daily_updates_ts_sum <- daily_updates_ts_sorted %>%  
  group_by(state,vote) %>% 
  summarise(mean_testing_rate = mean(Testing_Rate), med_testing_rate = median(Testing_Rate))
#model.2 <- lm(confirmed ~ c_d_rat, data = csse_today_states_sorted)
#summary(model.2)
```


Cases Time Series
```{r}

#matches("`(1?[0-9]{1})/([0-2]?[0-9]{1}/20)`")
cases_ts <- csse_confirmed_us_ts %>%
  select(-Lat, -Long_, -FIPS) %>%
  filter(Province_State %in% us_uid$Province_State) %>% 
  filter(!(is.na(Admin2))) %>% 
  filter(!(Province_State %in% c("Diamond Princess", "Recovered", "Grand Princess"))) %>%
  group_by(UID) %>%
  pivot_longer(`1/22/20`:last_col(), names_to = "date", values_to = "cases") %>%
  #gather(`1/22/20`:last_col(), key = "date", value = "cases") %>% 
  ungroup()
cases_ts  
cases_ts_states <- cases_ts %>%
  select(-Admin2) %>% 
  group_by(date, Province_State) %>%
  summarize(cases = sum(cases)) %>% 
  left_join(uid_states, key = "Province_State") %>%
  ungroup()

cases_ts_states %<>% 
  group_by(Province_State) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(cases_per_hundred_thousand = round((cases / Population * 100000), 1), cases_delta = cases - lag(cases), cases_relative = cases_delta / lag(cases))
cases_ts_states %<>%
  mutate(cases_delta = case_when(
    is.na(cases_delta) ~ as.double(cases),
    TRUE ~ as.double(cases) - as.double(lag(cases))
  ), cases_delta_per_hundred_thousand = round((100000 * cases_delta / Population), 1), cases_delta_ma = movingAverage(cases_delta), cases_delta_ma_per_hundred_thousand = movingAverage(cases_delta_per_hundred_thousand),
  cases_relative = case_when(
    is.na(cases_relative) ~ 0,
    cases_relative == Inf ~ 1,
    TRUE ~ cases_delta / lag(cases)), cases_perc = round(100*cases_relative, 3)) %>%
  ungroup() %>% 
  filter(!(date %in% exclude)) %>%
  mutate(state = factor(Province_State)) %>% 
  select(UID, date, state, Combined_Key, Population, cases, cases_per_hundred_thousand, cases_delta, cases_delta_per_hundred_thousand, cases_relative, cases_perc, cases_delta_ma, cases_delta_ma_per_hundred_thousand) %>% 
  mutate(vote = case_when(
  state %in% blue_states ~ "Dem",
  state %in% red_states ~ "Rep")) %>% 
  mutate(vote = factor(vote)) %>% 
  mutate(vote = fct_relevel(vote, c("Rep", "Dem")))
saveRDS(cases_ts_states, "cases_state.rds")

#cases_ts_states
```
Military Cases
```{r eval=FALSE, include=FALSE}
cases_ts_mil <- csse_confirmed_us_ts %>%
  select(-Lat, -Long_, -FIPS) %>%
  filter(Admin2 %in% branches)
cases_ts_mil
  group_by(UID) %>%
  gather(`1/22/20`:`4/22/20`, key = "date", value = "cases") %>% 
  ungroup()
cases_ts_mil
cases_ts_mil %<>%
  group_by(Admin2, date) %>% 
  summarize(cases = sum(cases)) %>%
  inner_join(military_uid, key = "Province_State") %>% 
  ungroup() %>% 
  group_by(Admin2) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(cases_per_hundred_thousand = round((cases / Population * 100000), 1), cases_delta = cases - lag(cases), cases_relative = cases_delta / lag(cases))
cases_ts_mil
cases_ts_mil %<>%
  mutate(cases_delta = case_when(
    is.na(cases_delta) ~ as.double(cases),
    TRUE ~ as.double(cases) - as.double(lag(cases))
  ), cases_delta_per_hundred_thousand = round((100000 * cases_delta / Population), 1), cases_delta_ma = movingAverage(cases_delta),
  cases_relative = case_when(
    is.na(cases_relative) ~ 0,
    cases_relative == Inf ~ 1,
    TRUE ~ cases_delta / lag(cases)), cases_perc = round(100*cases_relative, 3)) %>%
  ungroup()
saveRDS(cases_ts_mil, "cases_mil.rds")

```


Deaths Time Series
```{r}
deaths_ts <- csse_deaths_us_ts %>%
  select(-Lat, -Long_, -FIPS) %>%
  #filter(Province_State %in% states)# %>% 
  filter(!(is.na(Admin2))) %>% 
  filter(!(Province_State %in% c("Diamond Princess", "Recovered", "Grand Princess"))) %>%
  group_by(UID) %>%
  pivot_longer(`1/22/20`:last_col(), names_to = "date", values_to = "deaths") %>%
  ungroup()
deaths_ts_states <- deaths_ts %>%
  select(-Admin2) %>% 
  group_by(Province_State, date) %>% 
  summarize(deaths = sum(deaths)) %>%
  inner_join(uid_states, key = "Province_State", by="Province_State") %>%
  ungroup() %>% 
  group_by(Province_State) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(deaths_per_hundred_thousand = round((100000 * deaths / Population), 2), deaths_delta = deaths - lag(deaths), deaths_relative = deaths_delta / lag(deaths))
deaths_ts_states %<>%
  mutate(deaths_delta = case_when(
    is.na(deaths_delta) ~ as.double(deaths),
    TRUE ~ as.double(deaths) - as.double(lag(deaths))
  ), deaths_delta_per_hundred_thousand = round((100000 * deaths_delta / Population), 2), deaths_delta_ma = movingAverage(deaths_delta), deaths_delta_ma_per_hundred_thousand = movingAverage(deaths_delta_per_hundred_thousand),
  deaths_relative = case_when(
    is.na(deaths_relative) ~ 0,
    deaths_relative == Inf ~ 1,
    TRUE ~ deaths_delta / lag(deaths)), deaths_perc = round((100 * deaths_relative), 3), deaths_perc_ma = movingAverage(deaths_perc)) %>%
  ungroup() %>% 
  filter(!(date %in% exclude)) %>% 
  mutate(state = factor(Province_State)) %>% 
  mutate(vote = case_when(
  state %in% blue_states ~ "Dem",
  state %in% red_states ~ "Rep")) %>% 
  mutate(vote = factor(vote)) %>% 
  mutate(vote = fct_relevel(vote, c("Rep", "Dem")))
#deaths_ts_states <- deaths_ts_states %>% 
#  filter(!(is.na(state)), state %in% uid_states$Province_State)
saveRDS(deaths_ts_states, "deaths_state.rds")
#deaths_ts_states
```
Military Deaths
```{r eval=FALSE, include=FALSE}
deaths_ts_mil <- csse_deaths_us_ts %>%
  select(-Lat, -Long_, -FIPS) %>%
  filter(Province_State == "US Military") %>%
  filter(!(is.na(Admin2))) %>% 
  group_by(UID) %>%
  gather(`1/22/20`:`4/22/20`, key = "date", value = "deaths") %>% 
  ungroup()
deaths_ts_mil %<>%
  group_by(Admin2, date) %>% 
  summarize(deaths = sum(deaths)) %>%
  inner_join(military_uid, key = "UID") %>%
  ungroup() %>% 
  group_by(Admin2) %>%
  mutate(date = mdy(date)) %>%
  arrange(date) %>%
  mutate(deaths_per_hundred_thousand = round((100000 * deaths / Population), 2), deaths_delta = deaths - lag(deaths), deaths_relative = deaths_delta / lag(deaths))
deaths_ts %<>%
  mutate(deaths_delta = case_when(
    is.na(deaths_delta) ~ as.double(deaths),
    deaths_delta < 0 ~ (lead(deaths_delta) + lag(deaths_delta)) / 2,
    TRUE ~ as.double(deaths) - as.double(lag(deaths))
  ), deaths_delta_per_hundred_thousand = round((100000 * deaths_delta / Population), 2), deaths_delta_ma = movingAverage(deaths_delta),
  deaths_relative = case_when(
    is.na(deaths_relative) ~ 0,
    deaths_relative == Inf ~ 1,
    TRUE ~ deaths_delta / lag(deaths)), deaths_perc = round((100 * deaths_relative), 3), deaths_perc_ma = movingAverage(deaths_perc)) %>%
  ungroup()
saveRDS(deaths_ts_mil, "deaths_mil.rds")
```

Combined Tables
```{r message=FALSE}
combined_ts_states <- cases_ts_states %>% 
  full_join(deaths_ts_states) %>% 
  mutate(vote = case_when(
  state %in% blue_states ~ "Dem",
  state %in% red_states ~ "Rep")) %>% 
  mutate(vote = factor(vote)) %>% 
  mutate(vote = fct_relevel(vote, c("Rep", "Dem")))
#combined_ts_mil <- cases_ts_mil %>% 
#  full_join(deaths_ts_mil)
```

Top 5 states by cases
```{r}
top5cases_today <- cases_ts_states %>%
  filter(date == day) %>%
  top_n(25, cases_delta)

top5cases_today_pht <- cases_ts_states %>% 
  filter(date == day) %>% 
  top_n(25, cases_per_hundred_thousand)

top5deaths_today<- deaths_ts_states %>%
  filter(date == day) %>% 
  top_n(25, deaths_delta)

top5deaths_today_pht <- deaths_ts_states %>%
  filter(date == day) %>% 
  top_n(10, deaths_per_hundred_thousand)

priority_states <- c("Maryland", "Texas", "Virginia", "California", "Georgia", "Florida")

priority_cases_today <- cases_ts_states %>% 
  filter(date == day & (state %in% top5cases_today$state | state %in% priority_states)) %>% 
  mutate(state = fct_reorder(state, cases_delta))
priority_cases_today

priority_cases_today_pht <- cases_ts_states %>% 
  filter(date == day & (state %in% top5cases_today_pht$state | state %in% priority_states)) %>% 
  mutate(state = fct_reorder(state, cases_per_hundred_thousand))

priority_deaths_today <- deaths_ts_states %>% 
  filter(date == day & (state %in% top5deaths_today$state | state %in% priority_states)) %>% 
  mutate(state = fct_reorder(state, deaths_delta))

priority_deaths_today_pht <- deaths_ts_states %>% 
  filter(date == day & (state %in% top5deaths_today_pht$state | state %in% priority_states)) %>% 
  mutate(state = fct_reorder(state, deaths_per_hundred_thousand))
priority_deaths_today_pht


priority_ts <- combined_ts_states %>% 
  filter(state %in% top5cases_today$state | state %in% priority_states)
priority_ts_last30 <- priority_ts %>%
  filter(date %in% sinceMarch16)



```

New Cases Today in Priority States
```{r}
minor_values <- seq(as.integer(500), as.integer(round(max(priority_cases_today$cases_delta),-3)), by = as.integer(500))
p1 <- ggplot(priority_cases_today, aes(x = state, y = cases_delta, fill = vote)) +
  geom_col() +
  #scale_fill_manual(values = rep(blue.fill.10,3)) +
  labs(y = "", title = "New Cases", x = "")+
  theme(legend.position = "none",
        panel.grid.major.x=element_line(color="dark grey", size=0.2),
        panel.grid.minor.x=element_line(color="dark grey", size=0.2),
        axis.ticks=element_line(color="white"))+
  coord_flip() +
  #scale_y_discrete(breaks=as.character(c("2", "4", "5")))+#breaks=c(2000, 4000, 6000))+
  pq+jhu_caption
p1

```

New Cases Per 100,000 in Priority States
```{r}
p3 <- ggplot(priority_cases_today_pht, aes(x = state, y = cases_per_hundred_thousand, fill = vote)) +
  geom_col() +
  #scale_fill_manual(values = rep(blue.fill.10,3)) +
  labs(y = "", title = "Cumulative Cases (Per 100,000 People)", x = "") +
  theme(legend.position = "none",
        panel.grid.major.x=element_line(color="dark grey", size=0.2),
        panel.grid.minor.x=element_line(color="dark grey", size=0.2),
        axis.ticks=element_line(color="white")) +
  coord_flip()+
  pq+jhu_caption
p3
```

Deaths Today in Priority States
```{r}
p2 <- ggplot(priority_deaths_today, aes(x = state, y = deaths_delta, fill = vote)) +
  geom_col()+
  #scale_fill_manual(values = c(red.fill.10,red.fill.10)) +
  labs(y = "", title = "Deaths Today", x = "") +
  theme(legend.position = "none",
        panel.grid.major.x=element_line(color="dark grey", size=0.2),
        panel.grid.minor.x=element_line(color="dark grey", size=0.2),
        axis.ticks=element_line(color="white")) +
  coord_flip() +
  pq+jhu_caption
p2
```

Cumulative Deaths in Priority States
```{r}
p4 <- ggplot(priority_deaths_today_pht, aes(x = state, y = deaths_per_hundred_thousand, fill = vote)) + 
  geom_col() + 
  #scale_fill_manual(values = c(red.fill.10, red.fill.5)) + 
  labs(y = "", title = "Cumulative Deaths Per Hundred Thousand People", x = "") + 
  theme(legend.position = "none",
        panel.grid.major.x=element_line(color="dark grey", size=0.2),
        panel.grid.minor.x=element_line(color="dark grey", size=0.2),
        axis.ticks=element_line(color="white")) + 
  coord_flip() + 
  pq+jhu_caption
p4
```

```{r}
top3_states <- deaths_ts_states %>% 
  filter(date == day) %>% 
  top_n(7, deaths)
top3_states

#top3_states$state
top3_sum <- top3_states %>% 
  summarize(deaths = sum(deaths), state = "Top 3")
rest <- deaths_ts_states %>% 
  filter(!(state %in% top3_states$state)) %>%
  filter(date == day) %>% 
  summarize(deaths = sum(deaths), state = "Other")

compare <- full_join(rest, top3_sum)
#compare
#top3_sum$deaths/(rest$deaths + top3_sum$deaths)

p7 <- ggplot(compare, aes(x=state, y=deaths, fill = c("blue", "orange")))+
  geom_col() + 
  scale_fill_manual(values = c("blue", "orange")) + 
  labs(y = "", title = "Cumulative Deaths", x = "") + 
  theme(legend.position = "none") + 
  pq+jhu_caption
p7

red_states_ts <- combined_ts_states %>% 
  filter(date == day) %>% 
  filter(state %in% red_states) %>% 
  mutate(vote = "Rep")

blue_states_ts <- combined_ts_states %>% 
  filter(date == day) %>% 
  filter(state %in% blue_states) %>% 
  mutate(vote = "Dem")

red_v_blue <- full_join(red_states_ts, blue_states_ts)
red_v_blue %<>% mutate(vote = factor(vote)) %>% mutate(vote = fct_rev(vote))
#red_v_blue

red_v_blue_sum <- red_v_blue %>% 
  group_by(vote) %>% 
  summarise(average_death_rate = mean(deaths_per_hundred_thousand), average_case_rate=mean(cases_per_hundred_thousand), stdev=sd(cases_per_hundred_thousand))
#red_v_blue_sum

```
```{r}
p8 <- ggplot(red_v_blue, aes(x=deaths_per_hundred_thousand, fill = vote))+
  geom_density(alpha = 0.7)+
  #geom_segment(x=filter(red_v_blue_sum, vote=="Dem")$average_death_rate, xend=filter(red_v_blue_sum, vote=="Rep")$average_death_rate, y=0.0127, yend=0.013)+
  theme_538()
  #scale_fill_manual(values = c("blue", "red"))+
  #labs(x="Governor Party Affiliation", y="Death Rate")
p8
```

```{r}
#data.frame(a=seq(1,50000, by=1),b=rnorm(500000, mean=2500, sd=1000))
p9 <- ggplot(red_v_blue, aes(x=cases_per_hundred_thousand, fill = vote))+
  geom_density(alpha=0.7)+
  geom_density(data=data.frame(a=seq(1,500000, by=1),b=rnorm(500000, mean=mean(max(red_v_blue_sum$average_case_rate),min(red_v_blue_sum$average_case_rate)), sd=mean(max(red_v_blue_sum$stdev),min(red_v_blue_sum$stdev)))), aes(x=b, fill="Normal"), alpha=0.7)+
  #geom_segment(x=filter(red_v_blue_sum, vote=="Dem")$average_case_rate+250, xend=filter(red_v_blue_sum, vote=="Rep")$average_case_rate+500, y=5.37e-4, yend=4.3e-4)+
  theme(legend.position = "bottom")
  #scale_fill_manual(values = c("blue", "red"))+
  #labs(x="Governor Party Affiliation", y="Case Rate")+
  #theme_538()
p9
```


Cases Over Time (By State)
```{r}
 
xrange <-range(priority_ts_last30$date)
yrange <- range(priority_ts_last30$cases_delta_ma_per_hundred_thousand)

p5 <- ggplot(priority_ts_last30, aes(x = date, y = cases_delta_ma_per_hundred_thousand, color = state)) +
  geom_line(aes(group=state), size = 1) + 
  labs(y = "", title = "Daily New Cases per 100,000 Individuals", x = "Date") + 
  theme(legend.position = "bottom") + 
  annotate(geom = "text", x = xrange[1], y = yrange[2], label = paste(strwrap("Note: 7 Day Moving Average", 40), collapse = "\n"), hjust = 0, vjust = 1, size = 2)+ 
  scale_x_date(date_labels = "%b %d", date_breaks = "2 weeks")+
  facet_wrap("state", ncol = 1) +
  pq+jhu_caption
ggsave("11_states_cases.png", plot = p5, width = 8, height = 20, units = "in")
```

Daily Deaths per 10,000 People
```{r}
xrange <-range(priority_ts_last30$date)
yrange <- range(priority_ts_last30$deaths_delta_ma_per_hundred_thousand)

p6 <- ggplot(priority_ts_last30, aes(x = date, y = deaths_delta_ma_per_hundred_thousand, color = state)) +   geom_line(aes(group = state), size = 1)+ 
  #geom_smooth(aes(group = state), size = 1, se = FALSE, method = "loess", formula = 'y~x')+
  labs(y = "", title = "Daily New Deaths per 100,000 Individuals", x = "Date") + 
  theme(legend.position = "none") + 
  annotate(geom = "text", x = xrange[1], y = yrange[2], label = paste(strwrap("Note: 7 Day Moving Average", 40), collapse = "\n"), hjust = 0, vjust = 1, size = 2) + 
  facet_wrap("state", ncol = 2, shrink=FALSE) + 
  pq+jhu_caption
ggsave("11_states_deaths.png", plot = p6, width = 11, height = 8.5, units = "in")
```

Tree Map of Proportions
```{r}
tree_data <- red_v_blue %>% filter(date == max(date)) %>% top_n(25, deaths_per_hundred_thousand) %>% inner_join(uid)
tree_data
treemap(tree_data, index = "state", vSize = "Population", vColor="deaths_per_hundred_thousand", type = "dens", bg.labels = "transparent", fontcolor.labels=c("white"))

```

US National Statistics
```{r}
nation <- combined_ts_states %>% 
  arrange(date) %>% 
  group_by(date) %>% 
  filter(!(is.na(deaths))) %>% 
  summarise(population = us_uid$Population[5], cases = sum(cases), cases_it = cases / 1000, cases_delta = sum(cases_delta), cases_per_hundred_thousand = 100000 * cases / population, cases_perc = 100 * cases_delta / lag(cases), deaths = sum(deaths), deaths_it = deaths / 1000, deaths_delta = sum(deaths_delta), deaths_perc = as.double(100 * deaths_delta / lag(deaths)), deaths_per_hundred_thousand = 100000 * deaths / population) %>% 
  ungroup()

nation %<>% 
  mutate(lag_cases = as.double(lag(cases)), relative = cases_delta / lag_cases, cases_delta_ma = movingAverage(cases_delta),
  cases_perc = case_when(
    cases_perc == Inf ~ 100,
    TRUE ~ (100 * sum(cases_delta) / lag_cases)), cases_dir = (cases_delta_ma < lag(cases_delta_ma)),
  cases_perc_ma = movingAverage(cases_perc),
  lag_deaths = as.double(lag(deaths)), relative = deaths_delta / lag_deaths, deaths_delta_ma = movingAverage(deaths_delta),
  deaths_perc = case_when(
    deaths_perc == Inf ~ 100,
    TRUE ~ (100 * deaths_delta / lag_deaths)), deaths_dir = (deaths_delta_ma < lag(deaths_delta_ma)),
  deaths_perc_ma = movingAverage(deaths_perc)) %>% 
  ungroup()
nation[is.na(nation)] <- as.double(0)
nation[nation == Inf]<- as.double(100)
nation %<>%
  mutate(cases_dir = factor(nation$cases_dir, labels = c("red", "green")))
nation_last30 <- nation %>% 
  filter(date %in% sinceMarch16)
nation_today <- tail(nation_last30, 1) %>% select(-cases_it,-deaths_it,lag_cases,-relative,-cases_delta_ma,-cases_dir,-cases_perc_ma,-lag_deaths,-deaths_delta_ma,-deaths_dir,-deaths_perc_ma)
```


```{r}
typology <- data.frame(
  col_keys = c("", "Total", "24 Hour Change", "Total Per Hundred Thousand")
)

nation_today_cases <- nation_today %>% mutate(type="Confirmed Cases", cases_per_hundred_thousand = round(cases_per_hundred_thousand, 2)) %>% select(type, date, cases, cases_delta, cases_per_hundred_thousand) %>% rename(Date = date, "Total" = cases, "24 Hour Change" = cases_delta, "Total Per 100,000"= cases_per_hundred_thousand)

nation_today_deaths <- nation_today %>% mutate(type="Confirmed Deaths", deaths_per_hundred_thousand = round(deaths_per_hundred_thousand, 2)) %>% select(type, date, deaths, deaths_delta, deaths_per_hundred_thousand) %>% rename(Date = date, "Total" = deaths, "24 Hour Change" = deaths_delta, "Total Per 100,000"= deaths_per_hundred_thousand)

nation_today_table <- full_join(nation_today_cases, nation_today_deaths)
ft <- flextable(nation_today_table) 
ft <- autofit(ft)
ft
```

New Cases
```{r message=FALSE, warning=FALSE}
xrange <-range(nation_last30$date)
yrange <- range(nation_last30$cases_delta_ma)

nat_p1 <- nation_last30 %>% 
  ggplot(aes(x = date, y = cases_delta_ma)) + 
    geom_line(color = "blue", size = 1.2)+
    geom_area(fill="blue", alpha = 0.3)+
    #geom_smooth(aes(x=date, y=cases_delta), size=1, method="loess", se=FALSE)+
    #geom_rect(nation_last30, x=nation_last30$date, y = nation_last30$cases_delta_ma, mapping = aes(xmin = lag(date), xmax = date, fill = cases_dir), ymin = 0, ymax = Inf, alpha = 0.3) + 
    scale_color_manual(values = "blue")+
    scale_fill_manual(values = c("blue", "yellow"))+
    scale_x_date(date_breaks = "3 weeks", date_labels = "%b %d")+
    scale_y_continuous(breaks = breaks_extended(n=5), labels = scales::label_number_si(accuracy = 10), name = "Daily Cases", sec.axis = sec_axis(trans=~./nation_last30$population * 100000, name = "Per Hundred Thousand People"))+
    labs(y = "", title = "Daily New Cases", x = "Days") +
    theme(aspect.ratio = 6/14,
          legend.position = "none", 
          panel.grid.major.y=element_line(color="dark grey", size=0.1),
          panel.grid.major.x=element_line(color="dark grey", size=0.1),
          panel.grid.minor.y=element_line(color="dark grey", size=0.1),
          axis.ticks=element_line(color="black")) +
    pq+
    #ylim(0,max(nation_last30$cases_delta))+
    annotate(geom = "text", x = xrange[1], y = yrange[2], label = paste("Note: 7 Day Moving Average", collapse = "\n"), hjust = 0, vjust = 1, size = 4)+jhu_caption
  #First Wave Marker
  #geom_vline(xintercept=mdy("03-16-20"),color = "red", size = 1.2) + 
  #Second Wave Marker
  #geom_vline(xintercept=mdy("06-11-20"),color = "green", size = 1.2)+
  #Third Wave Marker
  #geom_vline(xintercept=mdy("09-12-20"),color = "purple", size = 1.2)


```

Percent Change in Deaths (5 Day Moving Average)
```{r warning = FALSE}
xrange <-range(nation_last30$date)
yrange <- range(nation_last30$deaths_delta_ma)

nat_p4 <- nation_last30 %>% ggplot(aes(x=date,y=deaths_delta_ma)) + 
    #geom_rect(nation_last30, x=nation_last30$date, y = nation_last30$deaths_delta_ma, mapping = aes(xmin = lag(date), xmax = date, fill = deaths_dir), ymin = 0, ymax = Inf, alpha = 0.3) + 
    geom_line(aes(x = date, y = deaths_delta_ma, color = ""), size = 1.2) +
    geom_area(fill="red", alpha = 0.3)+
    geom_area(data=nation_last30[nrow(nation_last30):nrow(nation_last30)-14,], mapping=aes(x=date, y=deaths_delta_ma), fill = "red")+
    scale_fill_manual(values = c("red", "green"))+
    scale_color_manual(values = red.fill.5)+
    scale_x_date(date_breaks = "3 weeks", date_labels = "%b %d")+
    scale_y_continuous(breaks = breaks_extended(n=6), labels = scales::label_number(accuracy = 1), name = "Daily Deaths", sec.axis = sec_axis(trans=~./nation_last30$population * 1000000, name = "Per Million People"))+
    labs(y = "", title = "Daily Deaths", x = "Date") +
    theme(aspect.ratio = 6/14,
          legend.position = "none", 
          panel.grid.major.y=element_line(color="dark grey", size=0.1),
          panel.grid.major.x=element_line(color="dark grey", size=0.1),
          panel.grid.minor.y=element_line(color="dark grey", size=0.1),
          axis.ticks=element_line(color="black")) +
    pq + 
    #ylim(0,max(nation_last30$deaths_delta_ma))+
    annotate(geom = "text", x = xrange[2], y = yrange[2], label = paste("Note: 7 Day Moving Average", collapse = "\n"), hjust = 1, vjust = 1, size = 4)+jhu_caption
nat_p1
nat_p4

```

Total Cases (in millions)
```{r, warning = FALSE}
nat_p2 <- nation_last30 %>% ggplot(aes(nation_last30, x=date, y=cases))+
  geom_line(size=1.1, color="#0016bd")+
  labs(title = "Cumulative Cases", x = "Date")+
  jhu_caption+
    theme(aspect.ratio = 6/14,
          legend.position = "none", 
          panel.grid.major.y=element_line(color="dark grey", size=0.1),
          panel.grid.major.x=element_line(color="dark grey", size=0.1),
          panel.grid.minor.y=element_line(color="dark grey", size=0.1),
          axis.ticks=element_line(color="black")) +
  scale_x_date(date_breaks = "3 weeks", date_labels = "%b %d")+
  scale_y_continuous(breaks = breaks_extended(n=10), labels = scales::label_number_si(accuracy = 1), name  = "Total", sec.axis = sec_axis(trans=~./(nation_last30$population) * 1000, name = "Per Thousand People"))+
  geom_area(alpha = 0.3, fill="blue")+
  geom_hline(yintercept=max(nation_last30$cases), color="red")+
  geom_hline(yintercept=nation_last30$cases[nrow(nation_last30)-14], color="green")+pq
nat_p2
```

Deaths Total
```{r, warning=F}
nat_p_deaths <- nation_last30 %>% ggplot(aes(nation_last30,x=date, y=deaths))+
  geom_line(size=1.1, color="#bf080e")+
  labs(title = "Cumulative Deaths", x = "Date")+
  jhu_caption+
    theme(aspect.ratio = 6/14,
          legend.position = "none", 
          panel.grid.major.y=element_line(color="dark grey", size=0.1),
          panel.grid.major.x=element_line(color="dark grey", size=0.1),
          panel.grid.minor.y=element_line(color="dark grey", size=0.1),
          axis.ticks=element_line(color="black")) +
  scale_y_continuous(breaks = breaks_extended(n=7), 
                     labels = scales::label_number_si(accuracy = 50),
                     name = "Total",
                     #sec.axis = dup_axis()
                     sec.axis=sec_axis(trans=~./(nation_last30$population) * 100000, name = "Per Hundred Thousand People"))+
  scale_x_date(date_breaks = "3 weeks", date_labels = "%b %d")+
  geom_area(fill="red", alpha = 0.5)+
  geom_hline(yintercept=max(nation_last30$deaths), color="red")+
  geom_hline(yintercept=nation_last30$deaths[nrow(nation_last30)-14], color="green")+pq
nat_p_deaths
```

Mortality Rate
```{r}
mortality_rate <- (nation$deaths-lag(nation$deaths, 7)) / (lag(nation$cases, 7) - lag(nation$cases, 14)) *100
mortality_rate[259]
```

Maps and Stuff
```{r}
#usstatesmap <- sf::st_read("maps/tl_2019_us_state.shp")
#st_drivers()
```